Challenges to transboundary fisheries management in North America under climate change

### 2020-12-14 Ecol/Manuscript_Draft.Rmd
Villy and Tay's comments

In 1982, the United Nations Law of the Sea Convention (UNCLOS) formalized the concept of exclusive economic zones (EEZs) creating what we know today as shared stocks [@UnitedNations:1986tl], i.e. stocks that migrate between countries EEZs (known as transboundary stocks) or between EEZs and the high seas (also called straddling stocks) [@Song:2017iua]. Today, an estimated of 347 [@Teh:2015gd] to 633 (Chapter 2) fish stocks cross national borders, some of them jointly managed by two or more countries. These stocks are responsible for more than 70% of these countries total fish catches (Chapter 2). Under Article 63, UNCLOS incentives actions to cooperate on the management of shared stocks [@UnitedNations:1986tl] as often management success depends on effective cooperation between parties [@Munro:2002uf; @Sumaila:2013hv]. Since the definition of shared stocks, game theory has been one of the most common approaches used to analyze the management of these type of stocks. However, shared stocks' management can be convoluted due to the participation of several fishing “players”, different countries and sometimes jurisdictions within a country, the migration patterns of the stock, and their abundance fluctuation within space and time [@Munro:2002uf, Engler and Saunders *in press*]. In addition, international treaties might not be prepared to address the effects that climate change will bring to shared fish stocks (Engler and Saunders *in press*, Koubrak and VanderZwaag, *in press*).

@@

The ocean is getting warmer [@Rheim:2013fv], less oxygenated [@Schmidtko:2017cl], and increasing in acidity [@Ross:2011kd; @IPCC:2019tn]. To cope with these changes in ocean biophysic properties, marine species, including shared fish stocks, have been shifting their distribution towards the poles and/or deeper waters [@Poloczanska:2016kk]. As climate change reshapes the ocean's environment worldwide [@Gattuso:2015jz], shared fisheries' delicate governance is threatened as new migration patterns may arise [@Miller:2013iv; @Pinsky:2018cb], historic distribution and abundances might shift [@Cheung:2010dt], and species basic natural traits may modify [@Pauly:2017hpa]. Catches of shared stocks like tunas, have significantly increased in some regions such as the subtropical Atlantic and western Pacific Oceans and are projected to continue [@MonllorHurtado:2017cm; @ErauskinExtramiana:2019fw]. Multiple shared species in North America have been observed to shift in distribution following changes in optimal conditions such as sockeye salmon (*Oncorhynchus nerka*) [@McDaniels:2010vn], Atlantic cod [@Pershing:2015gq], and flounders [@Pinsky:2012kq]. Moreover, these shifts are projected to continue towards the end of the $21^{st}$ century [@Cheung:2018hz]. As a result, some countries or management jurisdictions may see more shared fisheries and their catches shifting into their waters while others will stand to lose [@Oremus:2020id; @Pinsky:2018cb, Chapter 3]. Nevertheless, management rules for shared stocks (e.g., quota or spatial delimitation) are often determined based on current and/or historic knowledge of the stock's distribution and do not consider future shifts in distributions [@FredstonHermann:2018kp].

@@

The IPHC was established by Canada and the United States to oversee the management of Pacific halibut (*Hippoglossus stenolepis*) [@IPHC:2014tz]. There are 12 regulatory areas from which 3AB holds 51.2% of the stock, followed by regions 2ABC and 4ACDE with 23.1% and 20.4%, respectively, and lastly region 4B with only 5.2% of the stock distribution [@IPHC:2018vz]. In terms of management, the IPHC implements a total allowable catch (TAC) based on a yearly sampling of the Convention area in addition to a series of regulations to control fishing effort [@IPHC:2018vz]. The TAC is divided between recreational, subsistence and commercial fishery, with a portion set aside for bycatch of other fisheries [@IPHC:2019wt]. The commercial fishing season starts in March ending around November with restrictions allowing only set line gear with J-type hooks targeting individuals over 81.3 cm of total length [@IPHC:2019wt].

@@

Since 1998 Canada and the US have used a "Resource Sharing Understanding" to inform the management of Eastern George Bank's Atlantic cod (*Gadus morhua*) and haddock (*Melanogrammus aeglefinus*) and, yellowtail flounder (*Limanda ferruginea*) [@Pudden:2007bo; @TRAC:2016wu; @Song:2017va]. From 2010 onward, the GoMa suggests catch-limits based on a weighted method where 10% represents the stocks’ historical distribution (from 1967 to 1994) and 90% current distributions produced by quarterly surveys and catch [@TRAC:2016wu]. Since its introduction, the average quota allocation for each species proposed by the GoMa has been (Table \@ref(tab:ocTable1); Atlantic cod 77% Canada and 23% US, haddock 55% Canada and 45% US, and yellowtail flounder 34% Canada and 66% US [@TRAC:2015wj; @TRAC:2015vp; @TRAC:2015tt; @TRAC:2018tw; @TRAC:2018tj; @TRAC:2018wx; @Trinko:2019ou]. However, because this is an unofficial agreement, Canada and the US ultimately take single management decisions [@Soboil:2006gw]. In terms of management, the US has a multi species harvest control with area and season closures, mesh sizes, effort control, and mobile gear vessels that use bottom otter trawl gear [@Soboil:2006gw]. In contrast, Canada has a quota system in addition to limited-entry licensing, fleet allocations, and mesh and fish size regulation, among other input controls. Canada inshore vessels fish cod with longline and gillnet while haddock is mainly caught with bottom otter trawl gear [@Soboil:2006gw].

@@

Despite the expected decrease in MCP for the region, changes in the stock-share ratio of species within the Gulf of Maine show different outcomes dependent on the climate change scenario and species in question. Following a high emission path will affect mostly Canada’s share of yellowtail flounder and in less degree haddock, with an increase of cod share. under the low emission scenario, haddock and cod patterns intensify, while yellowtail flounder’s share approaches almost no change (Figure \@ref(fig:ocFigure5)). Such pattern is likely the combination of the bathymetry or the Gulf, the warming gradient, and the species distribution (see Discussion).

@@

The results of the present study suggest that climate change will alter the MCP of jointly managed transboundary fish stocks in North America consequently altering Canada’s and the US’s species’ stock-share ratio, regardless of the climate change scenario. These results are aligned with regional [@Morley:2018fn] projections suggesting that climate change will push marine species towards the poles and deeper water [@Pinsky:2013ab] in search of their ecological niche [@Poloczanska:2016kk]. Moreover, IPHC data[^IPHC] suggest that some of these shits are already happening. For example, since 2010, the distribution proportion of Pacific halibut has increased from 9% to 11% in region 2B, from 7.5% to 13% in region 2C, and from 12.3% to 13.5% in region 4CDE. On the other hand, regions 3A and 3B have seen the largest decreases in the IPHC regulatory areas since 2010, from 35.3% to 30.6% and 20.6% to 15.9%, respectively. Similarly, in the Gulf of Maine, the projected stock-share gain of yellowtail flounder and haddock by the US (Figure \@ref(fig:ocFigure5)) follows a historical trend where in 2019, Canada's stock-share decreased from 35% to 32% and 60% to 40% relative to 2010, respectively [@Trinko:2019ou].

@@

Geographic barriers [@Cheung:2015fm; @Kleisner:2016je], local temperature gradients [@Pinsky:2013ab], species interactions and human activities [@Serpetti:2017he] might change the rate and direction of species shifts. For the IPHC, geographic barriers might induce a westward increase of stock-share in IPHC regions where species can only migrate northward into the Arctic Ocean through the Bering Sea and Bering Strait [@Cheung:2015fm] (Figure \@ref(fig:ocFigure3)). In the Gulf of Maine, future projections could be a response to a temperature gradient shift combined with geographic barriers as southern waters are deeper and warming slower than northern waters according to the ESMs (Figure \@ref(fig:ocFigureS22)). Moreover, Maine has seen its landings of yellowtail flounder increased at the expenses of southern states [@Pinsky:2012kq]. This could be influencing the US gain in MCP in the GoMa in relation to Canada as species shift their distribution from lower latitudes naturally reaching the US (lower) region first. As the effects of climate change endure, even with high mitigation, joint plans should prepare to face changes in the stock-share ratio of transboundary stocks along both coast of North America.

The shifts in the distribution of transboundary stocks can jeopardize management objectives such as conservation measures and gear operation. Fish moving out of fishing grounds and into protected areas could result in a pressure increase to open such area to fishing. Moreover, overlapping shifting stocks could interfere in gear-limitation management rules of multiple fisheries generating conflicts between fleets [@VanDerVoo:2016uc]. The effectiveness of the IPHC-Closed Area ("CA" in Figure \@ref(fig:ocFigure2)) in terms of protecting juveniles has been historically questioned as trawling for other species is still allowed in the area [@Karim:2010ut; @IPHC:2017CA]. In 2015, for example, 97% of the trawl by-catch in areas 4CDE and the Closed Area were juveniles [@IPHC:2017CA]. Consequently, the Alaskan trawl fisheries has been closed before reaching annual quota due to the attainment of Pacific halibut bycatch quota limits [@Karim:2010ut]. Thus, the commission has been asked to open the closed area for Pacific halibut fishing, under the premise that the expansion of the trawl fishery is likely reducing any conservation goal for juvenile Pacific halibut [@IPHC:2017CA]. Although not assessed in this study, some trawling target species like Pacific cod (*Gadus macrocephalus*), flathead sole (*Hippoglossoides elassodon*), and Alaskan plaice (*Pleuronectes quadrituberculatus*) have already shift their distributions due to warming waters [@Stram:2009xc] and some are expected to continue shifting in similar direction than Pacific halibut [@Pinsky:2013ab]. The overlap of target species could be addressed by applying ecosystem-based management and dynamic management tools [@Hazen:2018fa] to manage these fisheries and reduce potential loss of sustainable harvest for both the halibut and the trawl fisheries.

@@

Side payments have been previously used to address changes in species distribution, including cases caused by environmental forcings. In game theory, a side payment is received by a player as a compensation from the other player in a shared resource agreement, with the premise that cooperation will result in a better overall outcome [@Bjorndal:2012vd; @Sumaila:2013hv]. Side payments do not have to be in monetary form and are widely used in transboundary stocks around the world. For example, Norway and Russia have implemented a quota swap strategy for jointly managed stocks of cod, haddock and capelin (*Mallotus villosus*) in the Barents Sea [@FAO:2020vw]. Similarly, species' quota swaps are allowed, up to a degree, within regulatory areas of the European union [@Baudron:2020dj]. Specifically, for northern European spring spawning herring (*Clupea harengus*), Norway, Iceland, Faroe Islands, Russia and the EU reached an agreement to manage the stock after its collapsed, partially due to climate variations [@Miller:2013iv]. Among the implemented rules, the agreement established a dynamic quota allocation, allowing members to fish part of their quota within Norway's EEZ, and land the catch in Norwegian ports. In North America, Canada and the US have previous history with the utilization of side payments when in the 70's Pacific salmon shifted its distribution resulting in large interceptions of Canada's salmon by Alaskan fisheries [@Miller:2013iv; @Song:2017va]. The conflict was resolved by the implementation of a conservation fund that worked as a side payment for both Canada and the state of Alaska [@Miller:2013iv; @Song:2017va]. The potential adaptation of side payments in terms of quota swaps or allocating EEZ-fishing rights across the Gulf of Maine EEZs could be a potential solution as stocks shift due to climate change.

Transboundary fisheries management have to be prepared for the uncertainty that comes with a changing world. Future climate change will depend on the path society as a whole will take, and thus I rely on scenario planning to account for the uncertainty build in future decision making [@vanVuuren:2011ef]. In our results, the "winners and losers" of climate change, and the intensity of the change, will be scenario dependent. For instance, stock-share of yellowtail flounder under a high emission scenario will be larger for the US while Canada's gain of cod stock-share will be larger under a low emission scenario (Figure \@ref(fig:ocFigure5)). Applying previously describe strategies (e.g., quota swaps or EEZ-fishing rights) could increase the resilience of treaties by preventing members from leaving the agreement due to a shift in threat points, as happened in the Pacific salmon case [@Miller:2013iv].

Models are attempts to represent reality (in our case a future reality) based on observational data, previously established theory, and future scenarios, and are thus, subjected to different degrees of uncertainty [@Payne:2016gn]. An ensemble of models is a way to present a more robust result that accounts for differences in the structural composition of each model [@Cheung:2016gv]. In here I used three ESMs to project future changes in species maximum catch potential. The levels of uncertainty related to the ESMs differ among case studies. Overall, results for the Gulf of Maine agree with a reduction in MCP of all three species. However, some discrete areas show a positive change for yellowtail flounder by mid $21^{st}$ century (Figure \@ref(fig:ocFigure3)), mainly driven by the GFDL model (Figure \@ref(fig:ocFigureS23)). Potential model artifacts could also be contributing to the results, especially in the northern part of the study area (Bay of Fundy) as most disagreeing grids are covered by land, which could be influencing the results. In contrast, considerable uncertainty exists in the change of MCP along the IPHC Convention area shown by a disagreement between ESMs (Figure \@ref(fig:ocFigureS24)). Off the coast of British Columbia, increasing temperature trends are consistent among ESMs, however, other processes such as acidification and deoxygenation are still not well understood from British Columbia to the Gulf of Alaska  [@TalloniAlvarez:2019gz]. Moreover, considerable uncertainty exists along the Bering Sea [@Douglas:2010Mb] and Antarctic Pacific regarding the extent and intensity of future sea-ice reduction under climate change [@Steiner:2015hc; @IPCC:2019tn]. Regarding the DBEM, its structural uncertainty has been previously tested for agreement against commonly used species distribution algorithms such as Maxent [@Phillips:2006ff] and AquaMaps [@Kaschner:2011hi; @Ready:2010wr] resulting in no qualitative differences in trends between algorithms [@Cheung:2016jd]. Finally, is worth mentioning that future changes to species distributions could be influenced by factors not captured by our model such as interactions between species [@Pecl:2017hu], adaptation of species to environmental changes, and anthropogenic factors [@Serpetti:2017he]. However, these factors are expected to increase the rate of range-shifting of the species making our results conservative [@Cheung:2010dt; @Serpetti:2017he].

>>>>

In 1982, the United Nations Law of the Sea Convention (UNCLOS) formalized the concept of exclusive economic zones (EEZs) creating the concept that we know today as shared stocks [@UnitedNations:1986tl], i.e. stocks that migrate between countries EEZs (known as transboundary stocks) or between EEZs and the high seas (also called straddling stocks) [@Song:2017iua]. Today, an estimated 347 [@Teh:2015gd] to 633 (Chapter 2) fish stocks have distributions that cross national borders, some of them jointly managed by two or more countries. These species are responsible for more than 70% of these countries total fish catches (Chapter 2). Under Article 63, UNCLOS incentives actions to cooperate on the management of shared stocks [@UnitedNations:1986tl] as management success often depends on effective cooperation between parties [@Munro:2002uf; @Sumaila:2013hv]. Since the definition of shared stocks, game theory has been one of the most common approaches used to analyze the management of these type of stocks. However, shared stocks' management can be convoluted due to the participation of several fishing “players”, different countries and sometimes jurisdictions within a country, the migration patterns of the stock, and their abundance fluctuation within space and time [@Munro:2002uf, Engler and Saunders *in press*]. In addition, international treaties might not be prepared to address the effects that climate change will bring to shared fish stocks (Engler and Saunders *in press*, Koubrak and VanderZwaag, *in press*).

@@

The ocean is getting warmer [@Rheim:2013fv], less oxygenated [@Schmidtko:2017cl], and increasing in acidity [@Ross:2011kd; @IPCC:2019tn]. To cope with these changes in ocean biophysical properties, marine species, including shared fish stocks, have been shifting their distribution towards the poles and/or deeper waters [@Poloczanska:2016kk]. As climate change reshapes the ocean's environment worldwide [@Gattuso:2015jz], shared fisheries' governance is threatened as new migration patterns may arise [@Miller:2013iv; @Pinsky:2018cb], historic distribution and abundances might shift [@Cheung:2010dt], and species basic natural traits may modify [@Pauly:2017hpa]. Catches of shared stocks like tunas, have significantly increased in some regions such as the subtropical Atlantic and western Pacific Oceans and are projected to continue [@MonllorHurtado:2017cm; @ErauskinExtramiana:2019fw]. Multiple shared species in North America have been observed to shift in distribution following changes in optimal conditions such as sockeye salmon (*Oncorhynchus nerka*) [@McDaniels:2010vn], Atlantic cod [@Pershing:2015gq], and flounders [@Pinsky:2012kq]. Moreover, these shifts are projected to continue towards the end of the $21^{st}$ century [@Cheung:2018hz]. As a result, some countries or management jurisdictions may see more shared fisheries and their catches shifting into their waters while others will stand to lose [@Oremus:2020id; @Pinsky:2018cb, Chapter 3]. Nevertheless, management rules for shared stocks (e.g., quota or spatial delimitation) are often determined based on current and/or historic knowledge of the stock's distribution and do not consider future shifts in distributions [@FredstonHermann:2018kp].

@@

The IPHC was established in 1923 by Canada and the United States to oversee the management of Pacific halibut (*Hippoglossus stenolepis*) [@IPHC:2014tz]. There are 12 regulatory areas from which 3AB holds 51.2% of the stock, followed by regions 2ABC and 4ACDE with 23.1% and 20.4%, respectively, and lastly region 4B with only 5.2% of the stock distribution [@IPHC:2018vz]. In terms of management, the IPHC implements a total allowable catch (TAC) based on a yearly sampling of the Convention area in addition to a series of regulations to control fishing effort [@IPHC:2018vz]. The TAC is divided between recreational, subsistence and commercial fishery, with a portion set aside for bycatch of other fisheries [@IPHC:2019wt]. The commercial fishing season starts in March ending around November with restrictions allowing only set line gear with J-type hooks targeting individuals over 81.3 cm of total length [@IPHC:2019wt].

@@

Since 1998 Canada and the US have used a "Resource Sharing Understanding" to inform the management of Eastern George Bank's Atlantic cod (*Gadus morhua*), haddock (*Melanogrammus aeglefinus*) and yellowtail flounder (*Limanda ferruginea*) [@Pudden:2007bo; @TRAC:2016wu; @Song:2017va]. From 2010 onward, the GoMa suggests catch-limits based on a weighted method where 10% represents the stocks’ historical distribution (from 1967 to 1994) and 90% current distributions as determined by quarterly surveys and catch [@TRAC:2016wu]. Since its introduction, the average quota allocation for each species proposed by the GoMa has been (Table \@ref(tab:ocTable1); Atlantic cod 77% Canada and 23% US, haddock 55% Canada and 45% US, and yellowtail flounder 34% Canada and 66% US [@TRAC:2015wj; @TRAC:2015vp; @TRAC:2015tt; @TRAC:2018tw; @TRAC:2018tj; @TRAC:2018wx; @Trinko:2019ou]. However, because this is an unofficial agreement, Canada and the US ultimately take single management decisions [@Soboil:2006gw]. In terms of management, the US has a multi species harvest control rule with area and season closures, mesh sizes, effort control, and mobile gear vessels that use bottom otter trawl gear [@Soboil:2006gw]. In contrast, Canada has a quota system in addition to limited-entry licensing, fleet allocations, and mesh and fish size regulation, among other input controls. Canada inshore vessels fish cod with longline and gillnet while haddock is mainly caught with bottom otter trawl gear [@Soboil:2006gw].

@@

Despite the expected decrease in MCP for the region, changes in the stock-share ratio of species within the Gulf of Maine show different outcomes dependent on the climate change scenario and species in question. Following a high emission path will negatively affect mostly Canada’s share of yellowtail flounder and in less degree haddock, with an increase of cod share. Under the low emission scenario, haddock and cod share patterns intensify, while yellowtail flounder’s share approaches almost no change (Figure \@ref(fig:ocFigure5)). Such pattern is likely due to the combination of the bathymetry or the Gulf, the warming gradient, and the species distribution (see Discussion).

@@

The results of the present study suggest that climate change will alter the MCP of jointly managed transboundary fish stocks in North America consequently altering Canada’s and the US’s species’ stock-share ratio, regardless of the climate change scenario. These results are aligned with regional [@Morley:2018fn] projections suggesting that climate change will push marine species towards the poles and deeper water [@Pinsky:2013ab] in search of their ecological niche [@Poloczanska:2016kk]. Moreover, IPHC data[^IPHC] suggest that some of these shifts are already happening. For example, since 2010, the distribution proportion of Pacific halibut has increased from 9% to 11% in region 2B, from 7.5% to 13% in region 2C, and from 12.3% to 13.5% in region 4CDE. On the other hand, regions 3A and 3B have seen the largest decreases in the IPHC regulatory areas since 2010, from 35.3% to 30.6% and 20.6% to 15.9%, respectively. Similarly, in the Gulf of Maine, the projected stock-share gain of yellowtail flounder and haddock by the US (Figure \@ref(fig:ocFigure5)) follows a historical trend where in 2019, Canada's stock-share decreased from 35% to 32% and 60% to 40% relative to 2010, respectively [@Trinko:2019ou].

@@

Geographic barriers [@Cheung:2015fm; @Kleisner:2016je], local temperature gradients [@Pinsky:2013ab], species interactions and human activities [@Serpetti:2017he] might change the rate and direction of species shifts. For the IPHC, geographic barriers might induce a westward increase of stock-share in IPHC regions where species can only migrate northward into the Arctic Ocean through the Bering Sea and Bering Strait [@Cheung:2015fm] (Figure \@ref(fig:ocFigure3)). In the Gulf of Maine, future projections could be a response to a temperature gradient shift combined with geographic barriers as southern waters are deeper and warming slower than northern waters according to the ESMs (Figure \@ref(fig:ocFigureS22)). Moreover, Maine has seen its landings of yellowtail flounder increase at the expenses of southern states [@Pinsky:2012kq]. This could be influencing the US gain in MCP in the GoMa in relation to Canada as species shift their distribution from lower latitudes naturally reaching the US (lower) region first. As the effects of climate change endure, even with high mitigation, joint plans should prepare to face changes in the stock-share ratio of transboundary stocks along both coast of North America.

The shifts in the distribution of transboundary stocks can jeopardize management objectives such as conservation measures and gear operation. Fish moving out of fishing grounds and into protected areas could result in a pressure increase to open such area to fishing. Moreover, overlapping shifting stocks could interfere in gear-limitation management rules of multiple fisheries generating conflicts between fleets [@VanDerVoo:2016uc]. The effectiveness of the IPHC-Closed Area ("CA" in Figure \@ref(fig:ocFigure2)) in terms of protecting juveniles has been historically questioned as trawling for other species is still allowed in the area [@Karim:2010ut; @IPHC:2017CA]. In 2015, for example, 97% of the trawl by-catch in areas 4CDE and the Closed Area were juveniles [@IPHC:2017CA]. Consequently, the Alaskan trawl fisheries has been closed before reaching annual quota due to the attainment of Pacific halibut bycatch quota limits [@Karim:2010ut]. Thus, the commission has been asked to open the closed area for Pacific halibut fishing, under the premise that the expansion of the trawl fishery is likely reducing any conservation goal for juvenile Pacific halibut [@IPHC:2017CA]. Although not assessed in this study, some trawling target species like Pacific cod (*Gadus macrocephalus*), flathead sole (*Hippoglossoides elassodon*), and Alaskan plaice (*Pleuronectes quadrituberculatus*) have already shifted their distributions due to warming waters [@Stram:2009xc] and some are expected to continue shifting in similar direction as Pacific halibut [@Pinsky:2013ab]. The overlap of target species could be addressed by applying ecosystem-based management and dynamic management tools [@Hazen:2018fa] to manage these fisheries and reduce potential loss of sustainable harvest for both the halibut and the trawl fisheries.

@@

Side payments have been previously used to address changes in species distribution, including cases caused by environmental forcing. In game theory, a side payment is received by a player as a compensation from the other player in a shared resource agreement, with the premise that cooperation will result in a better overall outcome [@Bjorndal:2012vd; @Sumaila:2013hv]. Side payments do not have to be in monetary form and are widely used in transboundary stocks around the world. For example, Norway and Russia have implemented a quota swap strategy for jointly managed stocks of cod, haddock and capelin (*Mallotus villosus*) in the Barents Sea [@FAO:2020vw]. Similarly, species' quota swaps are allowed, up to a degree, within regulatory areas of the European union [@Baudron:2020dj]. Specifically, for northern European spring spawning herring (*Clupea harengus*), Norway, Iceland, Faroe Islands, Russia and the EU reached an agreement to manage the stock after its collapsed, partially due to climate variations [@Miller:2013iv]. Among the implemented rules, the agreement established a dynamic quota allocation, allowing members to fish part of their quota within Norway's EEZ, and land the catch in Norwegian ports. In North America, Canada and the US have previous history with the utilization of side payments when in the 70's Pacific salmon shifted its distribution resulting in large interceptions of Canada's salmon by Alaskan fisheries [@Miller:2013iv; @Song:2017va]. The conflict was resolved by the implementation of a conservation fund that worked as a side payment for both Canada and the state of Alaska [@Miller:2013iv; @Song:2017va]. The potential adaptation of side payments in terms of quota swaps or allocating EEZ-fishing rights across the Gulf of Maine EEZs could be a potential solution as stocks shift due to climate change.

Transboundary fisheries management have to be prepared for the uncertainty that comes with a changing world. Future climate change will depend on the path society as a whole will take, and thus I rely on scenario planning to account for the uncertainty build in future decision making [@vanVuuren:2011ef]. In our results, the "winners and losers" of climate change, and the intensity of the change, will be scenario dependent. For instance, stock-share of yellowtail flounder under a high emission scenario will be larger for the US while Canada's gain of cod stock-share will be larger under a low emission scenario (Figure \@ref(fig:ocFigure5)). Applying previously described strategies (e.g., quota swaps or EEZ-fishing rights) could increase the resilience of treaties by preventing members from leaving the agreement due to a shift in threat points, as happened in the Pacific salmon case [@Miller:2013iv].

Models are attempts to represent reality (in our case a future reality) based on observational data, previously established theory, and future scenarios, and are thus, subjected to different degrees of uncertainty [@Payne:2016gn]. An ensemble of models is a way to present a more robust result that accounts for differences in the structural composition of each model [@Cheung:2016gv]. In here, I used three ESMs to project future changes in species maximum catch potential. The levels of uncertainty related to the ESMs differ among case studies. Overall, results for the Gulf of Maine agree with a reduction in MCP of all three species. However, some discrete areas show a positive change for yellowtail flounder by mid $21^{st}$ century (Figure \@ref(fig:ocFigure3)), mainly driven by the GFDL model (Figure \@ref(fig:ocFigureS23)). Potential model artifacts could also be contributing to the results, especially in the northern part of the study area (Bay of Fundy) as most disagreeing grids are covered by land, which could be influencing the results. In contrast, considerable uncertainty exists in the change of MCP along the IPHC Convention area shown by a disagreement between ESMs (Figure \@ref(fig:ocFigureS24)). Off the coast of British Columbia, increasing temperature trends are consistent among ESMs, however, other processes such as acidification and deoxygenation are still not well understood from British Columbia to the Gulf of Alaska  [@TalloniAlvarez:2019gz]. Moreover, considerable uncertainty exists along the Bering Sea [@Douglas:2010Mb] and Antarctic Pacific regarding the extent and intensity of future sea-ice reduction under climate change [@Steiner:2015hc; @IPCC:2019tn]. Regarding the DBEM, its structural uncertainty has been previously tested for agreement against commonly used species distribution algorithms such as Maxent [@Phillips:2006ff] and AquaMaps [@Kaschner:2011hi; @Ready:2010wr] resulting in no qualitative differences in trends between algorithms [@Cheung:2016jd]. Finally, is worth mentioning that future changes to species distributions could be influenced by factors not captured by our model such as interactions between species [@Pecl:2017hu], adaptation of species to environmental changes, and anthropogenic factors [@Serpetti:2017he]. However, these factors are expected to increase the rate of range-shifting of the species making our results conservative [@Cheung:2010dt; @Serpetti:2017he].
